00_datawrangling.Rmd

### 2019-12-19 analysis/00_datawrangling.Rmd
working on gene level counts not transcript levle

```{r countData}
# import count data, set rows at entreziz
countData <- read.table("../results/kallistocounts.txt", 
                        sep = ",", row.names = NULL)

## set row names as entrezid for deseq2
row.names(countData) <- countData$entrezid

# replace dashes with periods (to make deseq happy)
colnames(countData) <- gsub('-', '.', colnames(countData))

# create df with gene info and remove this from countData
geneinfo <- countData %>%
  select(row.names, Name, geneid, entrezid)
head(geneinfo)
countData <- countData %>%
  select (-c(row.names, Name, geneid, entrezid))

# because DESeq require integers
countData <- round(countData)

# print tolal num of genes and samples
dim(countData)

# make a tibble for the sole purpose of previewing this huge df
countDataTibble <- as.tbl(countData)
print(countDataTibble, n_extra = 0)
#rm(countDataTibble)
```

@@

write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)
```

>>>>

```{r countData}
# import count data, set rows at entreziz
kallistodata <- read.table("../results/kallistocounts.txt", 
                        sep = ",", row.names = NULL)

## save gene information
geneinfo <- kallistodata %>%
  select(row.names, Name, geneid, entrezid)
head(geneinfo)

# check for genes that have multple transcripts expressed
isoforms <- kallistodata %>%
  group_by(Name) %>%
  summarize(n = n()) %>%
  arrange(desc(n), Name) %>%
  filter(n > 1)
head(isoforms)

# for example, this gene has 
geneinfo %>% filter(Name == "CEND1")

# aggregate transcript counts to gene counts
countData <- kallistodata %>% 
  select(-row.names, -geneid, -entrezid) %>% 
  rename(gene = Name) %>%
  pivot_longer(-gene, names_to = "samples", values_to = "counts") %>%  
  pivot_wider(
    names_from = samples, 
    values_from = counts,
    values_fn = list(counts = sum))  %>% 
  arrange(gene)
head(countData)

countData <- as.data.frame(countData)
row.names(countData) <- countData$gene ## make gene the row name
countData[1] <- NULL ## make gene the row name
countData <- round(countData) #round all value to nearest 1s place

# replace dashes with periods (to make deseq happy)
colnames(countData) <- gsub('-', '.', colnames(countData))
head(countData,4)

# print tolal num of genes and samples
dim(countData)
```

@@

write.csv(geneinfo, "../metadata/00_geneinfo.csv", row.names = TRUE)

write.csv(isoforms, "../results/00_geneswithisoformss.csv", row.names = T)
```
